---
title: "Examples for inspiration"
author: "Markus Mortensen & André Waage Rivenæs"
date: "9 mars 2019"
output: 
 html_document:
   theme: cerulean
   toc: true
---

```{r setup, include = FALSE}
library(tidyverse)
library(lubridate)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Først leser vi inn data (merk: map_df dekkes i neste kurs)
```{r}
base_url <- "http://data.urbansharing.com/bergenbysykkel.no/trips/v1/2018/"
months <- seq(6, 12)
list_urls <- paste0(base_url, sprintf('%0.2d', months), ".csv")

df_raw <- map_df(list_urls, ~ read_csv(.x))

```

La oss legge til noen variabler i datasettet. 

```{r}
df <- df_raw %>%
  mutate(
    duration_hours = duration / 3600,
    duration_minutes = duration %/% 60,
    date = as_date(started_at),
    distance_trip = round(geosphere::distCosine(
      cbind(start_station_longitude, start_station_latitude),
      cbind(end_station_longitude, end_station_latitude)
    ))
  )
```


Deretter kan vi gjøre noen filtreringer ved å bruke dplyr::filter. Her lager vi et datasett med kun turer mellom 1000 og 3600 sekunder som hadde forskjellig start- og stoppsted, men hvor startstedet var Neumansgate.

```{r}
df_neumann <- df %>% 
  filter(duration > 1000 & duration < 3600) %>%
  filter(start_station_id != end_station_id) %>% 
  filter(start_station_name == "Neumannsgate")

```

Ved å bruke group_by + summarise kan vi f.eks. telle observasjoner per dato og finne snitt varighet per dato.

```{r}
df_neumann %>% 
  group_by(date) %>% 
  summarise(ant = n(),
            snitt_varighet = mean(duration)) %>% 
  arrange(-ant)
```

# Visualisering
## Scatter plots
```{r}
ggplot(data = df_neumann, aes(x = started_at, y = duration_hours)) + 
  geom_point()
```

Now we see we have some problems:

- The x-axis is broken
- Some points are overlapping
- It is really hard to spot a trend

Let's fix the input-data and use geom-jitter instead of geom point.

```{r}
ggplot(data = df_neumann, aes(x = as_date(started_at), y = duration_hours)) + 
  geom_jitter(alpha = 0.5)
  
```

We still don't really see what's going on here, so let's add a trend!

```{r}
ggplot(data = df_neumann, aes(x = as_date(started_at), y = duration_hours)) + 
  geom_jitter(alpha = 0.5, color = "grey") +
  geom_smooth()
```

Wow! Bike rides in the summer appear to have a longer duration than those in the fall/winter. 

Let's show the same relationship for the biggest start stations and do some cleaning:

```{r}
top_stations <- df %>% 
  count(start_station_name) %>% 
  top_n(9, n) %>% 
  pull(start_station_name)
  
df_top_stations <- df %>% 
  filter(start_station_name %in% top_stations) %>% 
  filter(duration_hours < 1)

ggplot(df_top_stations, aes(x = as_date(started_at), y = duration_hours)) +
  geom_jitter(alpha = 0.2, color = "grey") +
  geom_smooth() +
  facet_wrap(~start_station_name) +
  labs(x = "Date", y = "Duration") +
  theme_bw()
```


## Histograms and density plots

Let's create a simple histogram of the duration of bike rides.

```{r}
df <- df %>% 
  filter(duration_hours < 1)

df %>% 
  ggplot(aes(x = duration_minutes)) + 
  geom_histogram()
```

Now that's not very pretty, let's do some adjustments:
```{r}
df %>% 
  ggplot(aes(x = duration_minutes)) + 
  geom_histogram(fill = "steelblue", color = "black", bins = 60)
```

We can also make a density plot:

```{r}
df %>% 
  ggplot(aes(x = duration_hours)) + 
  geom_density(fill = "steelblue", color = "black")
```

We could split the density by a given variable.

```{r}

df <- df %>%
  mutate(is_longer_1km = distance_trip > 1000)

df %>% 
  ggplot(aes(x = duration_minutes, fill = is_longer_1km)) + 
  geom_density(color = "black", bins = 60, alpha = 0.6)
```

## Bar plot
Let's plot the average duration per station.

```{r}
summary_stations <- df %>%
  group_by(start_station_name) %>%
  summarise(avg_duration = mean(duration))

ggplot(summary_stations, aes(y = avg_duration, x = start_station_name)) +
  geom_bar(stat = "identity")
```

What a terrible plot! Let's fix it:

```{r}
ggplot(summary_stations, aes(y = avg_duration, x = fct_reorder(start_station_name, avg_duration))) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(x = "Station", y = "Average duration of each bike ride")
```

## Plotly
```{r}
library(plotly)

p <- df %>% 
  sample_n(100) %>%
  ggplot(aes(x = as_date(started_at), y = duration)) + 
  geom_point() + 
  geom_smooth()

ggplotly(p)
```


## Maps
```{r}
library(leaflet)

map_info <- df %>%
  group_by(start_station_name, start_station_longitude, start_station_latitude) %>%
  summarise(n_rides = n()) %>%
  ungroup() %>%
  mutate(n_rides_norm = 4 * n_rides / max(n_rides))

leaflet(map_info) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~ start_station_longitude, 
                   lat = ~ start_station_latitude, 
                   radius = ~ n_rides_norm, 
                   popup = ~ paste0(start_station_name, ", number of rides = ", n_rides),
                   fill = TRUE)
```

